An Axiomatic Basis for Computer Programming: Review

### 2017-02-16 hoare/paper.tex
Seems done with equation example

@@

Suppose we choose general quadratic formula for solving equation:
$$ x_{1,2} = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$$

@@

Suppose we didn't see that general formula is derived from equation. Lets just
apply substitution all occurrence of $x$ in postcondition by stated expressions.

@@

Lets predicate $B \equiv a \neq 0$, then precondition for our statement:
$$ P_1 \equiv P \wedge B $$

@@

With postcondition predicates, on the contrary, we couldn't make it
stronger.

Because if we don't change program $Q$, it will produce range of
results, which satisfy weaker predicate.

This range is superset of set which satisfy stronger predicate.

@@

This is actually Rule of Consequence stated in Hoare paper.

@@

In case when $x = 0$ (or $P \wedge \neg B$) equation takes the form:
$$ bx + c = 0 \text{, and } x = \frac{-c}{b}$$

Yet again this remains true when $C \equiv b \neq 0$ is true, so this statement
is correct according to precondition $P \wedge \neg B \wedge C$ and
postcondition $R$.

@@

In case when $b = 0$ or $\neg C$, equation takes the form:
$$ c = 0$$

Either $c = 0$ and equation holds for all $x \in \mathbb{R}$, or $c \neq 0$ and
equation

Actually we could replace whole program with one assignment:
$$ X = \emptyset $$

And it will be correct program, according to specification. That's why I
stressed that it doesn't help obtain right predicates.

>>>>

\usepackage{amsmath, amsfonts}
\renewcommand{\implies}{\Rightarrow}

@@

Suppose we choose general quadratic formula for solving equation:
\begin{equation}
 x_{1,2} = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
\end{equation}

@@

Suppose we didn't see that general formula is derived from equation. Lets just
apply substitution all occurrence of $x$ in postcondition by stated expression (1).

@@

Lets predicate $A \equiv a \neq 0$, then precondition for our statement:
$$ P_1 \equiv P \wedge A $$

@@

With postcondition predicates, on the contrary, we couldn't make it
stronger.
Because if we don't change program $Q$, it will produce range of
results, which satisfy weaker predicate.
This range is superset of set which satisfy stronger predicate.

@@

This is actually Rule of Consequence stated in Hoare paper.\\

@@

In case when $x = 0$ (or $P \wedge \neg A$) equation takes the form:
\begin{equation}
 bx + c = 0 \text{, and } x = \frac{-c}{b}
\end{equation}

Yet again, this remains true when $B \equiv b \neq 0$ is true, so this statement
is correct according to precondition $P \wedge \neg A \wedge B$ and
postcondition $R$.

@@

In case when $b = 0$ or $\neg B$, equation takes the form:
$$ c = 0$$

And again, two cases. Either $c = 0$ and equation holds for all
$x \in \mathbb{R}$:
\begin{equation}
  X = \mathbb{R}
\end{equation}

Or $c \neq 0$ and equation doesn't have solutions:
\begin{equation}
  X = \emptyset
\end{equation}

So lets combine this four statements and prove correctness of our program.

We will use Condition Rule, which could be derived from Rule of Iteration:
$$\frac{\{P \wedge A\} Q_1 \{R\}, \{P \wedge \neg A\} Q \{R\}}
{\{P\} \textbf{ if } A \textbf{ then } Q_1 \textbf{ else } Q \textbf{ end } \{R\}}$$

$Q_1$ correspond to equation (1), $Q_2$ to (2) and so on. $Q$ represent
nested statements:

Therefore, we prove that program holds postcondition $R$ for all values
satisfying $P$ and it is total correct:
\begin{align*}
  \{P\} & \textbf{ if } A \textbf{ then } Q_1 \textbf{ else } \\
        & \hspace{1em} \textbf{ if } B \textbf{ then } Q_2 \\
        & \hspace{2em} \textbf{ if } C \textbf{ then } Q_3 \textbf { else } Q_4 \textbf { end } \\
        & \hspace{1em} \textbf{ end }\\
        & \textbf{ end } \{R\}
\end{align*}

\rightline{{\rm $\Box$}}

Actually we could replace whole program with just one assignment:
$$ X = \emptyset $$

And prove it correctness according to $P$ and $R$ too, because universal
quantification expressed that every member of domain is satisfied some
predicate. This is true, when there is no members.

That's why I stressed that it doesn't help obtain right predicates.
